Drawing Trees and Animating Tree Changes

### 2010-05-12 paper/paper.tex
Small paper changes.

\usepackage{url}
\usepackage{graphicx}

@@

\author{Sandro Badame, Peter Boothe$^*$\\
 Manhattan College\\
\url{{sbadame.student,peter.boothe}@manhattan.edu}}

@@

\section{Introduction}
Trees, and especially trees which change over time, are fundamental objects in
computer science.  They occur in many disparate contexts, and while the uses of
these trees may be different, there are some commonalities which we can use to
build a tree visualization system which might be useful in many different
contexts.  In this paper we articulate the design principles of a general tree visualization system, present one such system that we have built in Java and made freely available, and then demonstrate our system's utility with several sample visualizations.

There are multiple reasons to visualize a system, and different reasons have
different requirements for the visualization.  A visualization designed for
experts may be very useful for that target group, but leave neophytes
confused.  Therefore, before we discuss the particulars of our visualization,
we must explain our reasons for wanting such a system in the first place, and
the expected audience.  We want to visualize dynamic trees for demonstration,
education, and beauty, and we would like our visualizations to be of benefit to
computer science undergraduates.

@@

It turns out that demonstrations are often useless.  But if we force students to build their own visualizations, then they actually learn.  So We would like to enable that as well.

@@

Past visualization frameworks went to great lengths to articulate the allowed
set of transformations which allowed them to heavily optimize their layout code
to minimize CPU usage.  In our situation, we would like to make our library as
usable as possible.  Therefore, we take every possible opportunity to make our
library easy to use, even if it increases the polynomial runtime of some of our
methods.  This is safe to do because we have restricted ourself to those graphs
which can be drawn, in their entirety, on a computer screen, and have therefore
restricted ourself to the (admittedly ill-defined) set of graphs which are
``small enough'' to still be drawn quickly using algorithms which may be of
order $\Omega(n^2)$, as long as these algorithms are not super-polynomial.

>>>>

\usepackage{url}
\usepackage{graphicx}
\usepackage[width=4.5in,height=7.2in]{geometry}

@@

\author{Sandro Badame, Peter Boothe\footnote{Corresponding author}\\
 Manhattan College\\
\url{{sbadame.student,peter.boothe}@manhattan.edu}}

@@

\section{Introduction}
Trees, and especially trees which change over time, are fundamental objects in
computer science.  They occur in many disparate contexts, and while the uses of
these trees may be different, there exist commonalities which we can use to
build a tree visualization system which might be useful in many different
contexts.  In this paper we articulate the design principles of a general tree visualization system, present one such system that we have built in Java and made freely available, and then demonstrate our system's utility with several sample visualizations.

There are multiple reasons to visualize a system, and different reasons have
different requirements.  A visualization designed for experts may be very
useful for that target group, but leave neophytes confused.  Therefore, before
we discuss the particulars of our visualization, we must explain our reasons
for wanting such a system in the first place, and the expected audience.  We
want to visualize dynamic trees for demonstration, education, and beauty, and
we would like our visualizations to be of benefit to computer science
undergraduates.

@@

It turns out that demonstrations are often useless.  But if we force students to build their own visualizations, then they actually learn.  So we would like to enable that as well.

@@

Past visualization frameworks went to great lengths to articulate the allowed
set of transformations which allowed them to heavily optimize their layout code
to minimize CPU usage.  In our situation, we would like to make our library as
usable as possible.  Therefore, we take every possible opportunity to make our
library easy to use, even if it increases the polynomial runtime of some of our
methods.  This is safe to do because we have restricted ourself to those graphs
which can be drawn, in their entirety, on a computer screen, and have therefore
restricted ourself to the (admittedly ill-defined) set of graphs which are
``small enough'' to still be drawn quickly using algorithms which may be of
order $\omega(n)$, as long as these algorithms are not super-polynomial.
